Combination of field and experimental data together with computational models

### 2022-02-09 manuscript_1.0/manuscript_1.0.Rmd
Changes addressing Zegni's comment in the ms

abstract: |
  While it is generally straightforward to quantify individual performance in cognitive experiments, identifying the underlying cognitive processes remains a major challenge. Often, different mechanistic underpinnings may yield similar performances, and Lloyd Morgan’s cannon warrants acceptance of the simpler explanation. Alternatively, when the different mechanisms interact with environmental conditions, variation in performance across environments might allow to statistically infer the mechanism responsible. However, to do this it is necessary to have quantitative predictions for the different candidate mechanisms. Here, we use a set of computational models to get quantitative predictions from alternative learning mechanisms, and using Bayesian statistics fit the model parameters to performance data. We used experimental data on performance in an ephemeral reward task by wild-caught cleaner fish *Labroides dimidiatus*, as well as cleaner and client fish densities from the locations of capture. The task can in principle be solved by estimating future consequences of an action, or by perceiving the removal of the ephemeral reward as psychological punishment (negative reinforcement). We found that a model where cleaners only estimate the future consequences of their actions explains best which cleaner relative abundances cause the fish to develop a preference for an ephemeral food source. This model also yields performances that can be considered the result of locally optimal decision-rules, in contrast to the negative reinforcement model.  We argue that the combination of computational models with data is a powerful tool to infer the mechanistic underpinning behind cognitive performance.

@@

Cooperation among individuals of the same or different species is
ubiquitous in nature. Whenever such interactions involve some degree 
of partner choice, biological market theory predicts that the exchange 
rate of goods and/or services traded adjusts to the law of supply and demand 
[@noe_Biological_1995a]. While biological market models have largely focused 
on evolutionary time scales [@noe_Biological_1994a;@johnstone_Mutualism_2008a] 
$\color{red}{\text{Hoekstra xxxx}}$, supply and demand conditions, which typically 
depend on the abundance of the species involved, certainly vary in time 
and space. Therefore, natural selection should 
favour the ability to flexibly adjust decisions and behavioural output to 
current market conditions. Indeed, such adjustments have been repeatedly 
documented [@axen_Signalling_1996]. In animals, an obvious general candidate 
mechanism for the strategic adjustment is the cognitive machinery. However, 
it is not clear which cognitive mechanisms allow individuals to adjust their 
behaviour to the varying market conditions. As adjustments to changes in 
market conditions have been documented in a great variety of species with 
highly variable brain anatomies, the question arises to what extent mechanisms 
beyond basic associative learning may be involved.

One example of strategic adjustment in a biological market is the marine
cleaning mutualism involving the cleaner fish *Labroides dimidiatus* and
'client' fish. Clients offer themselves as food patches for cleaners to
remove ectoparasites, which gives the cleaners calories and the clients
improved health [@waldie_LongTerm_2011;
@ros_Does_2011;@triki_Effects_2016]. As cleaners have about 2000
interactions per day [@grutter_Relationship_1995], they regularly face
the situation that two clients seek cleaning simultaneously. In such
conditions, it is important to distinguish between 'resident' and
'visitor' client species [@bshary_Cleaner_2001a]. The residents have
small home ranges and hence access to only the local cleaner, while the
visitors have large home ranges and hence access to two or more
cleaners. Thus, visitors have the option to switch to another cleaner if
being made to wait, while residents must wait for inspection. Indeed,
visitors have been observed to use their partner choice options in that
way [@bshary_Choosy_2002], which may explain why cleaners give
visitors service priority in a field study in the Red Sea
[@bshary_Cleaner_2001a]. More generally, cleaners apparently prefer to
interact with visitors [@triki_Marine_2021]. Such a preference may be
linked to a client's value as a food patch; visitors are on average
larger than residents [ @bshary_Cleaner_2001a], and size is positively
correlated with parasite load [@grutter_Relationship_1995]. However, 
early laboratory experiments showed that leaving causes service 
priority independent of client value [@bshary_Asymmetric_2002]. 
The experiments consisted of asimultaneous two-choice task, 
where cleaners had to decide which of two
Plexiglas plates, which offered equal amounts of food, to approach first
and eat mashed prawn of it. One plate was made to behave like a
resident, i.e. it would remain until the cleaner had eaten off it. The
other plate was made to behave like a visitor, i.e. it would be removed
if not inspected first. Cleaners learned to prefer the visitor plate and
hence obtained the double amount of food [@bshary_Asymmetric_2002].
Adult cleaner fish outperform various primates as well as rats and
pigeons in this original version of the biological market task
[@salwiczek_Adult_2012; @zentall_Early_2017]; only African grey
parrots solve this task as well [@pepperberg_Can_2014]. As the correct 
solution must be learned, and primates are known to readily distinguish 
between one and two reward units if presented in more conventional ways 
$\color{red}{\text{ref}}$, what makes the task difficult for most non-cleaners?

Recent research on the cognitive tool kit needed to solve an ephemeral 
reward task has used a broad approach as proposed by Shettleworth 
[@shettleworth_Cognition_2009], 
who defined cognition as including all ways in which animals take in information 
through the senses, process, retain and decide to act on it. Indeed, perception 
of relevant cues is of major importance for individual performance; 
i.e. what relevant information is encoded in the design of the two 
plates (colour versus size; [@wismer_Cuebased_2019]), or whether 
the information is in the colour of
the food [@pretot_Comparative_2021; @pretot_Comparing_2016], and
whether the food is visible or hidden/released only when the subjects
made choices on a touch screen [@pretot_Factors_2016]. However, identifying 
cues as salient is not the only challenge of the ephemeral reward task, 
as revealed by proximate learning models. Such models allow varying the 
cognitive tool kit and to evaluate which minimal kit is necessary to solve 
the task at hand (e.g. [@dubois_Model_2021]). Applied to the ephemeral 
reward task, learning models showed that basic reinforcement learning does 
not suffice to solve the task [@prat_Modelling_2022; @quinones_Reinforcement_2019]. 
This is particularly so when models assume the more complex natural 
situation in which cleaners not only face resident-visitor pairs but also 
visitor-visitor and resident-resident pairs, as well only a resident or 
only a visitor. To be able to give visitors priority over residents, 
cleaners need to be able to assess a client’s value separately for the 
three possible forms of encounter (alone, paired with a fish with the 
same strategic option, paired with a fish with the alternative strategic option) 
[@quinones_Reinforcement_2019]. The ability to keep such separate scores 
for the same stimulus has been termed configurational learning, 
or chunking, or segmentation (see references in [@prat_Modelling_2022]).

To solve the ephemeral reward task, cleaners also need to account for 
the future consequences of current decisions. In the model by 
Quiñones *et al.* [@quinones_Reinforcement_2019], this could be achieved in 
two non-mutually exclusive ways: through low temporal discounting of 
future effects, also termed ‘chaining’ [@enquist_Power_2016]; and/or 
through perceiving a visitor client leaving as psychological punishment 
(i.e. as a negative reinforcer). Given that a negative reinforcer is part 
of the ubiquitous learning mechanism termed operant conditioning 
[@thorndike_Animal_1898; @skinner_Behavior_1938], this explanation 
could be considered the simpler one. Furthermore, accounting for the 
future effects of an action is considered an advanced cognitive 
ability [@suddendorf_Evolution_2007]. Thus, standard logic of 
Lloyd Morgan’s cannon demands that operant conditioning as the simpler 
explanation is to be accepted by default. Here, we explore the 
possibility that fitting a model in which the action of these two 
mechanisms, operant conditioning and chaining, depends 
on model parameters may reveal which process cleaners actually use. 
This is because one process may be better at predicting the observed 
variation in cleaner fish performance as a function of ecological conditions.

Various repetitions of the ephemeral reward task in different years and 
with cleaners caught from various locations around Lizard Island, 
Great Barrier Reef, revealed that cleaner and client densities in 
nature affect how cleaners perform in the original market task: 
when relative densities are low, cleaners are more likely to fail to 
complete the task [@triki_Biological_2019;
@triki_Decrease_2018; @wismer_Variation_2014]. The explanation for this pattern
is twofold. First, low cleaner abundances implies low supply and high demand 
of cleaning services. Under this conditions, there will be less occasions under 
which failing to give priority to a visitor translates to an empty 
cleaning station. Secondly and also related to the state of the cleaning market,
visitor clients are less likely to exert their choice options 
(and hence wait for inspection if made to wait) when
cleaner fish densities are low [@triki_Brain_2020]. If visitors are not choosy, 
cleaners should not give them priority anymore. Thus, the underling 
cognitive machinery and ecological variables apparently interact to produce 
such variable performance. By fitting our model to the empirical data on 
fish densities and cleaner fish performance in the ephemeral reward task 
we aimed to address two questions. First, we aimed to determine the likely 
mechanism used by cleaners to incorporate future consequences of current 
decisions by testing whether operant conditioning, chaining, or a combination 
of both best explains performance. Second and related to the first 
question, we aimed to determine whether the two mechanisms differed 
with respect to the ecological conditions that are likely to cause high 
versus low performance in the ephemeral reward task. If that were the case, 
we would assess which mechanism would yield more appropriate performance 
patterns. Relying on the logic of biological market theory, we predicted 
that appropriate performance is to show low preference for visitors under high
local client-to-cleaner ratio as visitor leaving rates should be low 
and any leaving visitor will be readily replaced by another client. In contrast, 
cleaners should show high performance if the local client-to-cleaner 
ratio favours clients.

@@

To test the relative validity of chaining versus negative reinforcement as 
mechanisms to account for future consequences of current choices, 
we combined empirical data from Triki *et. al* 
[@triki_Biological_2019;@triki_Brain_2020]. with the computational 
model analysed in Quiñones *et al.*[@quinones_Reinforcement_2019]. In order to 
compare the explanatory power of the different hypothesis, we estimate the 
parameter values of three versions of the model that best fit the data. 
Particularly, we fit the parameter that determines how important 
future rewards are ($\gamma$), the magnitude of the negative reward 
($\eta$) and a scaling constant that translates absolute to relative 
abundances (see following sections for details). We obtained model 
predictions using a set of stochastic individual-based
simulations. Thus, each parameter combination of the model yields not one 
prediction for the data set, but a distribution of them. Thus, for the 
fitting procedure, we use one of the most common algorithms in bayesian inference,
the Markov Chain Monte Carlo (MCMC). MCMC is a stochastic algorithm that allows
to sample from the posterior distribution of model parameter values. The posterior
distribution represents the most likely parameter values given the data. 
Due to the stochastic nature of MCMC, which repeatedly samples parameter values
and evaluates their likelihood, the distribution of likelihoods from the 
stochastic model is also explored. This makes MCMC and bayesian inference
particularly suitable for the computational models we evaluate. Bellow, 
we explain first the computational model and the data collection, and later 
how the MCMC was used to estimate parameter values.

@@

The model consist of a set individual-based simulations where
individuals face a series of choices between two options. Individuals in
the model represent cleaners, and the options they face simulate the
natural conditions of a cleaning market. Cleaners experience a series 
of discrete time points in which they face different ‘states’, which 
are defined by the number and type of clients inviting for service. 
There are six possible states: zero clients, one resident, one visitor, 
two residents, two visitors, and a mix of one resident and one visitor. 
The probability of each state is largely determined by the relative 
abundance of cleaners, residents and visitors, but to some degree by a 
cleaner’s choices when it faces the resident-visitor combination. 
This is because residents are those clients that stay in the
cleaning queue when they are not given priority and hence would 
still be available in the next time point; while visitors leave
the queue (with a certain probability) after they are not given
priority. Individuals obtain a fixed reward from cleaning a client
regardless of the type. Every time individuals face and make a choice
they update the probability of making that same choice. The update is
based on the difference between the expectation of reward and the actual
reward the individual obtains (prediction error)
[@sutton_Reinforcement_2018; @rescorla_Theory_1972]; and it is carried
in the direction that would lead to more reward being obtained, given
the new information. In the long run, the probability of choosing a visitor 
over a resident converges in the model. To which probability the model 
will converge depends on the relative abundance of cleaners, visitors 
and resident; as well as on the probability of visitors leaving the station
when unattended. Further details of the model implementation can be
found in the paper [@quinones_Reinforcement_2019].

The model shows that only agents that calculate rewards separately for 
the different states (termed ‘fully aware agents’) can potentially 
prioritise visitors of residents in nature, and hence also solve the 
ephemeral reward task. This is because pooling rewards for the two 
clients from all states by ‘partially aware agents’ 
(that only recognize the two clients as different) cancels out 
all differences (see [@quinones_Reinforcement_2019]). However, 
such *chunking* by fully aware agents is not enough to develop a 
preference for visitors. In addition, agents need to find a mean 
to incorporate future consequences of current choices. In the model, 
this could be achieved with either of two parameters that could 
also work together.  First, $\gamma$ measures how
much individuals include future rewards in their decision updates. If
$\gamma=0$, individuals only use the immediate reward obtained from a
cleaning interaction. As $\gamma$ increases individuals include more the
reward obtained from the subsequent choices. That amounts to calculating
and using for decision making the future expected rewards of an action.
Second, $\eta$ measures how much individuals include in their reward the
fleeing behaviour of visitor clients as a negative component. Both of
these parameters allow individuals to use in their estimates the future
effects of their choices.

A further model analysis contrast the effect of these two 
parameters against variation in the relative abundance of cleaners 
and the two client types. Assuming certainty that visitors leave, the 
two parameters yield different 
predictions when relative client abundance is very high. Under such 
conditions, the analysis shows that under very high client (relative) 
abundance the presence of future rewards ($\gamma>0$) no longer favours 
the preference for the ephemeral clients, while the preference persists 
when the negative reward mediated by client behaviour ($\eta>0$) is 
included. Thus, the analysis suggests that future reward as an 
explanation for cleaner ability to prefer visitors is supported. 
At first sight, empirical data seem to support future rewards as 
the mechanism used by cleaners, as the statistical analysis  by 
Triki *et al.* [@triki_Biological_2019] data revealed that low 
cleaner-to-client ratios predict low preference for visitors and hence 
low performance in the ephemeral reward task. However, low cleaner-to-client 
ratios may also lead to low probabilities of visitors leaving, which would 
cause low preferences for visitors independently of the underlying 
mechanism [@quinones_Reinforcement_2019]. Furthermore, a combination 
of future and negative rewards, could still be yield a better explanation 
for the data than any single factor.

@@

The empirical data were collected between 2010 and 2018 always during 
the austral winter months June to August at Lizard Island 
($14.6682° S, 145.4604° E$), Great Barrier Reef, Australia. 
They consist of three sets: fish densities from a total of five study 
reef sites (Corner Beach-CB, Horseshoe-HS, 
Mermaid Cove-MC, Northern Horseshoe-NHS, and The Crest-TC), 
natural observations of cleaners to quantify the probability 
of visitors leaving if initially ignored in favour of another client, and the 
performance of individuals caught from these sites in the ephemeral reward test. 
In total, we have twelve site/year data sets for fish densities and 
corresponding performance in the test. Thus, some sites were sampled more 
than once. Due to major environmental perturbations during the study period 
($\color{red}{\text{refs}}$), fish densities changed significantly even 
within sites [@triki_Decrease_2018;@triki_Fluctuations_2019]. To estimate 
the population density of cleaner fish and their clients at a given site in 
a given year, Triki *et al.* [@triki_Biological_2019] used a series of ten 
transects of $30m$ each. Observers swam along the transect lines placed on 
the reef and first counted the visible large-bodied adult fish 
(species with total length TL > $10cm$) on a width of $5m$, and then on the 
return individuals of small-bodied fish species (TL < $10 cm$) on a width 
of $1 m$. We included cleaner fish in the $5 m$ transect. Based on the 
classification by Bshary [@bshary_Cleaner_2001a] and Wismer *et al.* 
[@wismer_Variation_2014], all visitor client species ($n = 124$) were 
large-bodied. Resident clients comprised all small-bodied species 
($n = 30$) but also some large-bodied ones ($n = 15$). Overall, 
large-bodied species are highly correlated with being a visitor, 
and small-bodied species are a good correlate of residents. Based on the 
transect data, Triki *et al.* [@triki_Biological_2019;@triki_Brain_2020] 
estimated cleaner fish, small-bodied client, large-bodied client and total 
client fish densities per $100 m^2$. LASSO analyses revealed that cleaner and
large-bodied client densities were highly correlated, and only the former 
was hence kept in the analyses as representative of both [@triki_Biological_2019]. 
Cleaner densities, in interaction with visitor client leaving probability, 
explained best individual performance in the ephemeral reward task 
[@triki_Biological_2019], while cleaner-to-client ratio was not a significant 
factor. This creates a certain mismatch with the model by Quiñones *et al.* 
[@quinones_Reinforcement_2019], which focused on abundance of cleaners relative 
to overall client abundance in order to capture the characteristics 
of the market. Importantly, relative abundances of clients define not only 
the probability of interactions with residents and visitors, but also affect 
how often the cleaning station is empty (e.g. there are no clients to be cleaned). 
The higher the relative abundance of cleaners is, the higher is the probability 
of cleaning stations being empty at a given time point. How often the 
cleaning stations are empty in turns influences the value that cleaners can 
expect from their choices. The emptier the cleaning stations are, the more 
beneficial it is to prioritize a client that will not stay. On the other hand, 
situations where cleaners have to choose will be rare if cleaner relative 
abundance is high, which reduces the total value of prioritizing visitors 
over residents and makes it more difficult to learn a preference for visitors. 
Given the model structure, we used the empirical data to calculate relative 
abundances of cleaners, residents and visitors for the current analyses. 
To incorporate the empirical finding that absolute densities matter, we 
introduced a new scaling constant (see section on model fitting).

>>>>

abstract: |
  While it is generally straightforward to quantify individual performance in cognitive experiments, identifying the underlying cognitive processes remains a major challenge. Often, different mechanistic underpinnings yield similar performances, and Lloyd Morgan’s cannon warrants acceptance of the simpler explanation. Alternatively, when the different mechanisms interact with environmental conditions, variation in performance across environments might allow to statistically infer the mechanism responsible. However, to do this it is necessary to have quantitative predictions for the different candidate mechanisms. Here, we use a set of computational models to get quantitative predictions from alternative learning mechanisms, and using Bayesian statistics fit the model parameters to performance data. We used experimental data on performance in an ephemeral reward task by wild-caught cleaner fish *Labroides dimidiatus*, as well as cleaner and client fish densities from the locations of capture. The task can in principle be solved by estimating future consequences of an action, or by perceiving the removal of the ephemeral reward as psychological punishment (negative reinforcement). We found that a model where cleaners only estimate the future consequences of their actions explains best which cleaner relative abundances cause the fish to develop a preference for an ephemeral food source. This model also yields performances that can be considered the result of locally optimal decision-rules, in contrast to the negative reinforcement model.  We argue that the combination of computational models with data is a powerful tool to infer the mechanistic underpinning behind cognitive performance.

@@

Disentangling the mechanistic underpinnings of behaviour 
is typically done under controlled conditions in a laboratory setting. However, 
controlled laboratory conditions often mask the inter and intra-specific 
variation that arises in the interaction between mechanisms and 
environmental variation. From a evolutionary perspective, mechanisms 
are likely selected because of how they allow individuals to respond to 
environmental variation. For example, biological market theory predicts that 
the exchange rate of goods and/or services traded between cooperative 
partners adjusts to the law of supply and demand, when individuals have 
some degree of partner choice [@noe_Biological_1995a]. 
Supply and demand conditions, which typically 
depend on the abundance of the species involved, certainly vary in time 
and space. Therefore, natural selection should 
favour the ability to flexibly adjust decisions and behavioural output to 
current market conditions. Indeed, such adjustments have been repeatedly 
documented [@axen_Signalling_1996]. In animals, an obvious general candidate 
mechanism for the strategic adjustment is the cognitive machinery. However, 
it is not clear which cognitive mechanisms allow individuals to adjust their 
behaviour to the varying conditions. As adjustments to changes in 
market conditions have been documented in a great variety of species with 
highly variable brain anatomies, the question arises to what extent mechanisms 
beyond basic associative learning may be involved.

One example of strategic adjustment in a biological market is the marine
cleaning mutualism involving the cleaner fish *Labroides dimidiatus* and
'client' fish. Client fish seek cleaner fish services at their territory 
(so-called “cleaning station”) and offer themselves as food patches 
to get their ectoparasites removed, which provides cleaners 
with food and clients with improved health [@waldie_LongTerm_2011;
@ros_Does_2011;@triki_Effects_2016;@demaire_Reduced_2020]. 
Given the capacity of some client fish to swim larger distances and 
access multiple cleaning stations while others access the only cleaning 
station at their territory, it is crucial to categorize clients as 
either “visitors” or “residents”, respectively. During cleaning interactions, 
a cleaner fish often face a choice between a visitor and a resident client 
seeking its cleaning services simultaneously. Visitors have the option 
to switch to another cleaner fish if being made to wait, while residents 
must wait for inspection. Indeed,
visitors have been observed to use their partner choice option in that
way [@bshary_Choosy_2002], which may explain why cleaners give
visitors service priority in a field study in the Red Sea
[@bshary_Cleaner_2001a].

When researchers aimed at testing wild-caught cleaner fish, as 
well as individuals from other species, abilities 
to prioritize visitor over resident clients in a lab-based paradigm, 
they used Plexiglas plates of different colours and shapes offering same 
amount of food as surrogates for visitor and resident clients.
One plate is made to behave like a resident, i.e. it remains until 
the cleaner had eaten off it. The other plate was made to behave 
like a visitor, i.e. it is removed if not inspected first.
Cleaners learned to prefer the visitor plate and
hence obtained the double amount of food [@bshary_Asymmetric_2002].
Furthermore, adult cleaner fish outperformed various primates as well as rats and
pigeons in this original version of the biological market task
[@salwiczek_Adult_2012; @zentall_Early_2017]; yet the African grey
parrots solve this task as well [@pepperberg_Can_2014]. Given that primates 
are known to readily distinguish 
between one and two reward units if presented in more conventional ways 
$\color{red}{\text{ref}}$, this raises the question:
what makes the task difficult for most non-cleaners?

Recent research on the cognitive tool kit needed to solve the ephemeral 
reward task has used a broad approach as proposed by Shettleworth 
[@shettleworth_Cognition_2009], who defined cognition as including 
all ways in which animals acquire information 
through the senses, process, retain and decide to act on it. 
Indeed, perception of relevant cues is of major importance 
for individual performance. For example, some animals find relevant 
information in the plates 
 [@wismer_Cuebased_2019] while others find it in the food 
[@pretot_Comparative_2021; @pretot_Comparing_2016]. However, identifying 
cues as salient is not the only challenge of the ephemeral reward task, 
as revealed by proximate learning models. Such models allow varying the 
cognitive tool kit and to evaluate which minimal kit is necessary to solve 
the task at hand (e.g. [@dubois_Model_2021]). Applied to the ephemeral 
reward task, learning models showed that basic reinforcement learning does 
not suffice to solve the ephemeral reward task [@prat_Modelling_2022;
@quinones_Reinforcement_2019]. 
This is particularly so when models assume the more complex natural 
situation in which cleaner fish not only face resident-visitor pairs but also 
visitor-visitor and resident-resident pairs, as well a resident alone or 
a visitor alone. To be able to give visitors priority over residents, 
cleaners need to be able to assess a client’s value separately for the 
three possible scenarios (alone, paired with a fish with the 
same strategic option, paired with a fish with the alternative strategic option) 
[@quinones_Reinforcement_2019]. The ability to distinguish and value differently
one stimulus alone from compound versions of it has been termed 
configurational learning, or chunking, 
or segmentation (see references in [@prat_Modelling_2022]).

To solve the ephemeral reward task, cleaners need to account for 
the future consequences of current decisions. In the model by 
Quiñones *et al.* [@quinones_Reinforcement_2019], this could be achieved in 
two non-mutually exclusive ways: through low temporal discounting of 
future effects, also termed ‘chaining’ [@enquist_Power_2016]; and/or 
through perceiving a visitor client leaving as psychological punishment 
(i.e. as a negative reinforcer). Low temporal discounting is when 
individuals include in their valuation of an action the reward effects 
that this will have in the future. This is done by combining in a single 
valuation the reward obtained in the current time with all the reward that 
comes after, discounting for how far in the future reward 
is accrued. 'Chaining' the reward of these different time steps allows 
individuals to take actions that increase the long-term reward at the sacrifice
of short term considerations [@enquist_Power_2016]. Even though, 'chaining' 
can be readily implemented computationally in learning models 
[@enquist_Power_2016;@sutton_Reinforcement_2018], cognitively it seems to be 
a complex adaptation [@suddendorf_Evolution_2007]. On the other hand,
using client behaviour as a negative reinforcer is in principle 
easier to implement. Negative reinforcement is part of the ubiquitous 
learning mechanism termed operant conditioning [@thorndike_Animal_1898; 
@skinner_Behavior_1938]. Thus, standard logic of 
Lloyd Morgan’s cannon demands that operant conditioning as the simpler 
explanation is to be accepted by default. Ideally, the two mechanisms should 
be evaluated in light of how well they explain the available data.

Over the last decade, our team tested over a hundred of wild-caught 
cleaner fish in the exact same paradigm of  of the ephemeral reward 
task [@salwiczek_Adult_2012;@wismer_Variation_2014;@triki_Decrease_2018;
@triki_Biological_2019;@triki_Brain_2020]. These fish often come from 
different reef locations around the study field site at Lizard Island, 
Great Barrier Reef in Australia. Further investigation of the local 
eco-sociological conditions revealed that cleaner and client 
fish population densities have a substantial impact on cleaner fish 
performance in the task. Cleaner fish from reef sites with relatively 
low densities were more likely to fail at solving the task [@triki_Biological_2019;
@triki_Decrease_2018; @wismer_Variation_2014]. The explanation for this pattern
is twofold. First, low cleaner fish population densities imply 
low supply and high demand for cleaning services. 
Under such conditions, there are fewer occasions under 
which failing to give priority to a visitor translates to an empty 
cleaning station. Second and also related to the state of the cleaning market,
visitor clients at low cleaner fish population densities 
are less likely to exert their partner choice options 
and are hence more likely to wait for inspection if not given cleaning 
priority [@triki_Brain_2020]. In the absence of partner choice 
cleaners should not give priority to visitors any more.

The interplay between the underling cognitive machinery and 
local ecological conditions apparently generate the documented variable 
performance among individuals of the same species. 
Our approach of fitting the computational model to the empirical data on 
fish densities and cleaner fish censuses and performance
in the ephemeral reward task aimed at: (i), determining which 
mechanism cleaner fish use to incorporate future consequences of current 
decisions by testing whether operant conditioning, chaining, or a combination 
of both best explains their performance; (ii)  determining whether the 
two mechanisms differed with respect to the ecological conditions 
that are likely to cause high versus low performance in the 
ephemeral reward task. Additionally, we assessed which mechanism 
yields optimal performance patterns. Relying on the logic of 
biological market theory, we predicted 
that appropriate performance is to show low preference for visitors under low
local cleaner-to-client ratio. 
<!-- In this case, as visitor leaving rates  -->
<!-- should be low as well and any leaving visitor will be readily  -->
<!-- replaced by another client in the queue. In contrast,  -->
<!-- cleaners should show high performance in the ephemeral reward task -->
<!-- if the local cleaner-to-client ratio favours clients.  -->

@@

In this study, we combined empirical data from Triki *et. al* 
[@triki_Biological_2019;@triki_Brain_2020] and the computational 
model analysed in Quiñones *et al.*[@quinones_Reinforcement_2019] 
to test the relative validity of chaining versus operant conditioning as 
mechanisms to account for future consequences of cleaner fish desicion-making.
In order to compare the explanatory power of the different hypotheses, 
we estimated the parameter values of three versions of 
the model that best fit the empirical data. 
Particularly, we fitted the parameter that determines how important 
future reward is ($\gamma$), the magnitude of the negative reward 
($\eta$) and a scaling constant that translates cleaner fish absolute 
to relative population abundances (see following sections for details). 
We obtained the model predictions using a set of stochastic individual-based
simulations. Thus, each parameter combination of the model yielded not one 
but a distribution of predictions. For the 
fitting procedure, we used the bayesian inference algorithm: 
Markov Chain Monte Carlo (MCMC). MCMC is a stochastic algorithm that allows
to sample from the posterior distribution of model parameter values. The posterior
distribution represents the most likely parameter values given the data. 
Due to the stochastic nature of MCMC, which repeatedly samples parameter values
and evaluates their likelihood, the distribution of likelihoods from the 
stochastic model is also explored. This makes MCMC and bayesian inference
particularly suitable for the computational models we evaluate. Bellow, 
we explain first the computational model and the data collection, and later 
how the MCMC was used to estimate parameter values.

@@

The model consists of a set individual-based simulations where
individuals, representing cleaners, face a series of choices 
between two options, which simulate the natural conditions 
of the cleaning market. Individuals experience a series 
of discrete time points in which they face different ‘states’, 
defined by the number and category of client fish (visitor or residents) 
inviting for cleaning services. There are six possible states: 
zero clients, one resident, one visitor, 
resident-resident, visitor-visitor, and resident-visitor. 
The probability of each state is largely determined by the relative 
abundance of cleaner fish, residents and visitors, but to some degree by
cleaner fish choices when it faces the resident-visitor combination. 
This is because residents are willing to queue for cleaning service; 
while visitors leave the queue (with a certain probability) when made to wait. 
Individuals obtain a fixed reward from cleaning a client fish
regardless of the category. Every time individuals face and make a choice
they update the probability of making that same choice. The update is
based on the difference between the expected reward and the obtained
reward (prediction error) [@sutton_Reinforcement_2018; 
@rescorla_Theory_1972]. Given the new information, the update is carried
in the direction that leads to more reward being obtained, given
the new information. In the long run, the probability of choosing a visitor 
over a resident converges in the model. To which probability the model 
will converge depends on the relative abundance of cleaners, visitors 
and resident; as well as on the probability of visitors leaving the 
cleaning station when made to wait. Further details of the model 
implementation can be found in Quiñones *et al* [@quinones_Reinforcement_2019].

The model shows that agents need to find a mean 
to incorporate future consequences of current choices. In the model, 
this could be achieved with either of two parameters that could 
also work together.  First, $\gamma$ measures how
much individuals include future rewards in their decision updates. If
$\gamma=0$, individuals only use the immediate reward obtained from a
cleaning interaction. As $\gamma$ increases individuals include more the
reward obtained from the subsequent choices. That amounts to calculating
and using for decision making the future expected rewards of an action.
Second, $\eta$ measures how much individuals include in their reward the
fleeing behaviour of visitor as a negative component. Both of
these parameters allow individuals to use in their estimates the future
effects of their choices.

<!-- A further model analysis contrast the effect of these two  -->
<!-- parameters against variation in the relative abundance of cleaners  -->
<!-- and the two client types. Assuming certainty that visitors leave, the  -->
<!-- two parameters yield different  -->
<!-- predictions when relative client abundance is very high. Under such  -->
<!-- conditions, the analysis shows that under very high client (relative)  -->
<!-- abundance the presence of future rewards ($\gamma>0$) no longer favours  -->
<!-- the preference for the ephemeral clients, while the preference persists  -->
<!-- when the negative reward mediated by client behaviour ($\eta>0$) is  -->
<!-- included. Thus, the analysis suggests that future reward as an  -->
<!-- explanation for cleaner ability to prefer visitors is supported.  -->
<!-- At first sight, empirical data seem to support future rewards as  -->
<!-- the mechanism used by cleaners, as the statistical analysis  by  -->
<!-- Triki *et al.* [@triki_Biological_2019] data revealed that low  -->
<!-- cleaner-to-client ratios predict low preference for visitors and hence  -->
<!-- low performance in the ephemeral reward task. However, low cleaner-to-client  -->
<!-- ratios may also lead to low probabilities of visitors leaving, which would  -->
<!-- cause low preferences for visitors independently of the underlying  -->
<!-- mechanism [@quinones_Reinforcement_2019]. Furthermore, a combination  -->
<!-- of future and negative rewards, could still be yield a better explanation  -->
<!-- for the data than any single factor.  -->

@@

The empirical data were collected between 2010 and 2019 always during 
the austral winter months June to August from a total of five 
study reef sites (Corner Beach-CB, Horseshoe-HS, Mermaid Cove-MC, 
Northern Horseshoe-NHS, and The Crest-TC)  at Lizard Island 
($14.6682° S, 145.4604° E$), Great Barrier Reef, Australia. 
The data consist of three sets: fish censuses, field 
observations of cleaner-client interactions to quantify the probability 
of visitors leaving if made to wait, and the 
performance of wild-caught cleaner fish in the ephemeral reward test. 
In total, we have twelve site/year data sets for fish censuses and 
corresponding performance in the lab test. Thus, some sites were sampled more 
than once. To estimate the population density of cleaner fish 
and their clients at a given site in 
a given year, Triki *et al.* [@triki_Biological_2019] used a series of ten 
transects of $30m$ each. Observers swam along the transect lines placed on 
the reef and first counted the visible large-bodied adult fish 
(species with total length $TL \geq 10cm$) including cleaner fish 
on a width of $5m$, and then on the 
return individuals of small-bodied fish species ($TL < 10 cm$) \
on a width of $1 m$
(see Triki et al. 2018 for further details on fish censuses data collection). 
We then scaled the counts of cleaner fish, small-bodied, 
and  large-bodied  clients fish densities per $100 m^2$. 
From the study by Triki *et al* [@triki_Biological_2019] 
cleaner fish and large-bodied client populations densities were 
highly correlated, and only the former 
was hence used in the analyses as representative of both measures.

<!-- Cleaner densities, in interaction with visitor client leaving probability,  -->
<!-- explained best individual performance in the ephemeral reward task  -->
<!-- [@triki_Biological_2019], while cleaner-to-client ratio was not a significant  -->
<!-- factor. This creates a certain mismatch with the model by Quiñones *et al.*  -->
<!-- [@quinones_Reinforcement_2019], which focused on abundance of cleaners relative  -->
<!-- to overall client abundance in order to capture the characteristics  -->
<!-- of the market. Importantly, relative abundances of clients define not only  -->
<!-- the probability of interactions with residents and visitors, but also affect  -->
<!-- how often the cleaning station is empty (e.g. there are no clients to be cleaned).  -->
<!-- The higher the relative abundance of cleaners is, the higher is the probability  -->
<!-- of cleaning stations being empty at a given time point. How often the  -->
<!-- cleaning stations are empty in turns influences the value that cleaners can  -->
<!-- expect from their choices. The emptier the cleaning stations are, the more  -->
<!-- beneficial it is to prioritize a client that will not stay. On the other hand,  -->
<!-- situations where cleaners have to choose will be rare if cleaner relative  -->
<!-- abundance is high, which reduces the total value of prioritizing visitors  -->
<!-- over residents and makes it more difficult to learn a preference for visitors.  -->
<!-- Given the model structure, we used the empirical data to calculate relative  -->
<!-- abundances of cleaners, residents and visitors for the current analyses.  -->
<!-- To incorporate the empirical finding that absolute densities matter, we  -->
<!-- introduced a new scaling constant (see section on model fitting).  -->

The field observation data consisted of video recordings/encodings of the 
cleaner-client cleaning interactions. There were videos from eight 
cleaners per site/year of a duration of $30 min$ each. Triki *et al* 
extracted information from every event wherein a visitor client 
was made to wait in favour of another client (visitor or resident), and noted 
whether or not the visitor left or queued for the cleaning service
[@triki_Biological_2019;@triki_Brain_2020].
